---
title: "ggplot2"
author: "Sabarish"
date: "17 June 2016"
output: html_document
---

# Graphs and visualisations with R

R has a very powerful package called ggplot2 developed by Mr Hadley Wickham who is the chief scientist at R Studio. He has also created other packages like dplyr which I will discuss in a separate blogpost. 
His book ggplot2 is available here - ggplot2 book by Hadley Wickham

So, first let us discuss why are graphs useful?
Mainly because, you don't need to look at the entire data to arrive at a judgement about the data. This saves you a lot of time. Imagine having 10000 rows of data for 10 years of revenues of your company!
It allows you to understand the overall trend of the data without actually looking at the data at all.
It allows you a good summary of the current situation of the data and aids better understanding of the data.

So what is ggplot2 and how is it different from inbuilt graphics of R?

ggplot2 is built on the principles of grammar of graphics. It basically builds the graphic layer by layer.
Think of it as the composition of your salad: While making the salad, first you add the veggies, then the dressing, then spices etc.
Similarly, in ggplot2; you first add what you want to plot known as the aesthetics, then you add the title of the graph, then the title of the axes etc.
But remember, it is very well possible to eat your salad without the dressing or spices as per you preferences. Similarly, you can have a plot without the title and the axes names. That is the beauty of ggplot2, you can simply have whatever you want!

Below are some of the basic codes in R.
First I will call the library ggplot2. 
If you haven't installed it yet, you can install it on your R Studio by going to the window where you see Files, Plots, Packages, Help and Viewer and there search for ggplot2 in you Packages.

Or simply run the following command on your console:
install.packages("ggplot2", dependencies = TRUE)    
```{r}
library(ggplot2)
```   
There is an inbuilt data set in R called the diamonds dataset. For more information:    
```{r}
?diamonds
```   

Lets now observe the quality of cut versus the number of diamonds that belong to a certain quality.
So I call the ggplot function, enter name of the dataset. For the aesthetics, I give what I want on the x axis. When I don't specify what I want on the y axis, it by default takes as the count of whatever is there on the x axis, just like any other graphic tool. Now the '+' function indicates what sort of a graph I want. There are many options but here I will use the bar chart.   
```{r}
ggplot(diamonds, aes(x=cut))+geom_bar()
```   

Here are few other examples:
```{r}
ggplot(diamonds, aes(x=x, y = price))+geom_point()
ggplot(diamonds, aes(x=y, y = price))+geom_point()
ggplot(diamonds, aes(x=z, y = price))+geom_point()
```   

Here is another small dataset which is useful for plotting:   
```{r}
?BOD
```   
Here I show you how to post the axes labels and the title to the plot   
```{r}
ggplot(BOD,aes(x=as.factor(Time),y=demand))+geom_bar(stat="identity",col = "red", fill = 10)+xlab("Time")+ylab("Demand")+ggtitle("Demand as a function of Time")
```   
Going back to the Diamonds dataset:   
```{r}
ggplot(diamonds,aes(x=cut,fill = clarity))+geom_bar(col = "black",position = "dodge")
```   
Here's another dataset for top hitters in 2001 in MLB season available in the gcookbook library   
```{r}
library(gcookbook)
ggplot(tophitters2001[1:25,],aes(x=avg,y=name))+geom_point()
```   
You can also flip coordinates by using the flipping function    
```{r}
library(gcookbook)
ggplot(tophitters2001[1:25,],aes(x=avg,y=name))+geom_point()+coord_flip()
```   
Another dataset is the US Population change dataset in the library gcookbook.
Suppose we want to colour to change as per the region of colour. Since it is a variable we put it inside the aesthetic rather than in geom bar
manually change the colour of the bars.
This is as complex a plot as it can get. But look at the syntax. It is so easy. Everything between the two '+' signs are separate commands for the data.    
```{r}
library(gcookbook)
?uspopchange
ggplot(uspopchange[1:25,],aes(x=reorder(Abb,Change),y=Change,fill=Region))+geom_bar(stat="Identity")+scale_fill_manual(values=c("Orange","yellow", 3, 10))+ggtitle("US Population Change")  